Brief Announcement: An Improved Lower Bound for Dynamic Balanced Graph Partitioning

### 2020-02-17 paper/paper.tex
wip
preliminaries

\section{Problem Definition}
\label{sec:problem-definition}

Formally, the online \emph{Balanced RePartitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
We say that nodes in $C \subset V$ are \emph{collocated}
% in a partition $P$
if they are assigned to the same cluster.
%\maciek{We can ommit the formal definition: "I.e.,
%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". %However, if we use the notion of cluster later, then TODO: define cluster}
%\maciek{We can ommit the formal definition: "I.e.,
%$\forall u,v \in C: \mathit{cluster} (u,P) = \mathit{cluster} (v,P)$.". However, if we use the notion of cluster later, then TODO: define cluster}
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Before serving the request,
the online algorithm can re-partition,
i.e.,
it may (optionally) move some of the nodes into clusters different than their current assigned clusters.
Once the re-partition is executed,
we serve
the communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.

\section{Preliminaries}

\maciek{if preliminaries are short, we should merge them with the previous section}

An algorithm that  keeps all components collocated is \emph{component-respecting}.

To this end, we maintain \emph{components of nodes}, in the similar fashion to previous approaches\cite{repartition-disc}.
We say that \TAlg is \emph{component respecting}, i.e., it maintains the invariant that nodes of each component are located in the same cluster.

We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.

This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
such that frequently communicating nodes (from different clusters)
are relocated to the same cluster before they inflict too much communication cost.
A \emph{reconfiguration} (i.e., re-partitioning) is performed by
moving (i.e., migrating) nodes between clusters at a cost.
The goal is to jointly minimize communication and migration costs while serving the sequence.

@@

  \maciek{Components are not introduced. Thus, the reader does not know what a single node means.}
	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
	Next,
	we issue a request between $x_0$ and $y_0$.
	Since this is an external \maciek{external request is not defined at this point} request,
	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
	For this,
	\ALG{} moves to a new configuration
	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.

@@

%We define the algorithm \TAlg in the following way.
Now we describe the algorithm \TAlg.
It partitions nodes into components, and
intially each node belongs to its own singleton component.
For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component respecting partition.
If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
\maciek{Serving remotely: say explicitly}

@@

    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
    Note that $U\neq V$, as otherwise the request would not trigger a~reconfiguration.
    We consider cases upon types of clusters $U$ and $V$.
    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, and costs at most $3\alpha$.
    \maciek{at most $2\alpha$ is sufficient}

@@

  As \TAlg is component respecting, it never increases any counter above $\alpha$.
  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.

>>>>

\section{Preliminaries}
\label{sec:problem-definition}

The online \emph{Balanced Re-Partitioning} problem (\OBRP{}) is defined as
follows. We are given a set of $n$ nodes,
initially arbitrarily partitioned into $\ell$~clusters,
each of size~$k$.
The input consists of a sequence of pairwise communication requests
$\sigma = (u_1,v_1),$ $(u_2,v_2),$ $(u_3,v_3), \ldots$,
where a pair $(u_t,v_t)$ indicates that nodes $u_t$ and $v_t$ exchange a fixed amount of data.
Nodes in $C \subset V$ are \emph{collocated}
if they reside in the same cluster.
An algorithm serves a communication request between two nodes
either \emph{locally} at cost~0
if they are collocated,
 or \emph{remotely} at cost~1
 if they are located in different clusters.
 We refer to the two types of requests as \emph{internal}
 and \emph{external} requests, respectively.
Before serving a request,
an online algorithm can \emph{re-partition},
i.e.,
it may (optionally) move some nodes into clusters different than their current assigneclusters while respecting the capacity of every cluster. he re-partition is executed,
the algorithm serves the  request
the communication request between two  nodes at cost~0 if they are collocated,
or at cost~1 if they are located in different clusters.
The cost of migrating a node from one cluster to another
is~$\alpha \in \mathbb{Z}^+$.
For any algorithm $ALG$,
its cost,
denoted by $ALG(\sigma)$,
is the total cost of communications and
the cost of migrations performed by $ALG$ while serving the sequence $\sigma$.

 useful sentences. This section is TODO}
An algorithm that  keeps all components collocated is \emph{component-respecting}.
%
A \emph{component} is a subset of frequently communicating nodes.   ly,                                                   ,every node is trivially a \emph{singleton} component.
We grow components similar fashion to previous approaches~\cite{repartition-disc}.
which proved to be an effective techique for mimi                                                                it always keep.
That is,
if the algorithm needs to move a node,
it moves the whole component containing the node to the sameemph{component -respecting}, i.e., it maintains the invariant that nodes of each component are collocated in the same cluster.
%We say that the request $(u,v)$ is \emph{external}, if the algorithm is in the configuration, where $u$ and $v$ are located at different clusters.
%
This is achieved by opting a sequence of  partitions (a.k.a.~configurations)
such that frequently communicating nodes (from different clusters)
are relocated to the same cluster before they inflict too much communication cost.
A \emph{reconfiguration} (i.e.,\re-partitioning) is performed by
 migrating) nodes betwen clusters at a cost.   ,s to jointly minimize communication and migration costs while serving the sequence.

@@

  \maciek{Components are not introduced. Thus, the reader does not know what a single node means.}
	Let $x_0$  be the only single node left on $S_1$ and  $y_0 \in S_2$ be any single node on $S_2$.
	Next,
	we issue a request between $x_0$ and $y_0$.
	Since this is an external request,
	\ALG{} joins them into one component and collocates them in some cluster other than $S_1$.
	For this,
	\ALG{} moves to a new configuration
	that replaces $x_0$ and $y_0$ with two other single nodes $x_1$ and $y_1$ respectively.

@@

%We define the algorithm \TAlg in the following way.
Now we describe the algorithm \TAlg.
It partitions nodes into components, and
intially each node belongs to its own singleton component.
For each pair of nodes, \TAlg maintains a counter, initiated to $0$. 
Upon receiving a request to a pair that is not collocated at one cluster, it increases their counter by $1$.
If the counter for a pair $(u,v)$ reaches $\alpha$, \TAlg merges the components of $u$ and $v$, and moves to the closest component-respecting partition.
If no such partition exists, \TAlg resets all components to singletons, resets all counters to $0$, and ends the phase.
\maciek{Serving remotely: say explicitly}

@@

    Consider a request between $u$ and $v$ that triggered the repartition and let $U$ and $V$ be their respective clusters.
    Note that $U\neq V$,
	 as otherwise the request would not trigger a~re-partitioning.
	 We consider cases upon types of clusters $U$ and $V$.
    Note that this is impossible that either $U$ or $V$ is of type $C_3$, as otherwise we merge components of size $3$, and no component-respecting partition exists, a contradiction.
    If either $U$ or $V$ is of type $C_1$, then this cluster can fit the merged component, and the rebalance is local within $U$ and $V$, and costs at most $3\alpha$.
    \maciek{at most $2\alpha$ is sufficient}

@@

  As \TAlg is component-respecting, it never increases any counter above $\alpha$.
  We say that the pair $(u, v)$ is \emph{saturated}, if the counter has value $\alpha$, and \emph{unsaturated} otherwise.
